Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis

### 2022-12-06 SALT-abstract/SALT-abstract.tex
removed tex gyre termes font
abstract updates 12/6

%% fonts
\setmainfont[Scale=MatchLowercase,Mapping=tex-text,SmallCapsFont={TeX Gyre Termes}, SmallCapsFeatures={Letters=SmallCaps}]{Times New Roman}
\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Times New Roman}

@@

\vspace{-.4\baselineskip}
\noindent We present experimental evidence that the projectivity of attitude complements varies across different entailment-cancelling operators, and that this effect of entailment-cancelling operator differs between various attitude verb triggers. However, our findings about the interaction of operator and trigger do not support \posscite{karttunen_observations_1971} classification of factive vs. semi-factive verbs.

@@

	\vspace{-.4\baselineskip}
	Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one, affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}. In light of this, we expect that the hetergeneous entailment-cancelling operators in \ref{ex:family} affect projection differentially.

	\citet{karttunen_observations_1971} proposed generalizations to this effect, distinguishing between \emph{factive} verbs (\emph{regret, forget, resent}) and \emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}), suggesting that factives always project, while semi-factives always project across negation, but not always in polar questions or conditionals. To provide a systematic way of distinguishing these classes, \cite{djarv_cognitive_2018} suggest that they correspond to emotive and cognitive predicates, respectively. Regarding the effect of operators on projection, \cite{smith_relationship_2014} find an effect of .. on ... (?) Here, we experimentally address the questions: \textbf{(i)} Is the projection of content affected by differences in entailment-canceling environments? \textbf{(ii)} Do these effects vary for different verbs (and in what way)?

@@

\vspace{-\baselineskip}
\paragraph{Method.} \hspace{-1em}
	The study consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause. Experiments 1--3 used polar question embedding, exps. 4--6 used negation, 7--9 modals, and 10--12 conditionals, making the \texttt{operator} manipulation a between-subjects factor. We analyzed data from 2682 self-identified native English speakers participated online across the 12 experiments (recruited via Prolific and Amazon MTurk). Participants saw items Latin-squared and randomized with six control stimuli.

@@

\paragraph{Results and Analysis.} \hspace{-1em}
	Pooling the data across our 12 experiments, we examined the effect on certainty-ratings of \texttt{operator}, and \texttt{verb}, as well as their interaction. Mean certainty-ratings by operator and predicate, and 95\% bootstrapped confidence intervals are shown in  \textbf{Figure \ref{fig:figure1}}.

	\begin{figure}[h]
		\vspace{-.8\baselineskip}
		\centering
		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
		\caption{\small Mean certainty ratings by predicate and operator}
		\label{fig:figure1}
	\end{figure}

	\vspace{-.3\baselineskip}
	% \textbf{version w baseline know/m}
	% \noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{know} and \texttt{modal} as reference levels, and random intercepts for participants and items. We found highly significant main effects of \texttt{operator} (\texttt{conditional}: $+0.142$, \texttt{negation}: $+0.143$, ; \texttt{question}: $+0.225$, where $p < 2e^{-16}$ in all three cases), as well as many interactions of \texttt{operator} and  \texttt{verb} across the board (where $p < 0.001$ in $38$ cases, $p < 0.01$ in one, and $p < 0.05$ in three out of $57$ possible interactions).\\

	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items. The mean for this baseline (intercept) is $0.867$. We found highly significant main effects of \texttt{operator} for \texttt{conditional} ($-0.116$, $p < 1.6e^{-13}$) and \texttt{modal} ($-0.156$, $p < 2e^{-16}$), while the effect of \texttt{question} was only marginally significant ($+0.025$, $p < 0.1$). We also found many interactions of \texttt{operator} and \texttt{verb} across the board (where $p < 0.001$ in $43$ cases, $p < 0.01$ in three, and $p < 0.05$ in one out of $57$ possible interactions).\\

	Notably, the interactions of:
	\begin{itemize}
		\item discover and question, conditional: significant positive interaction effects (can we really argue that for discover: $Q, C > N$, based on this? I am not really sure how to do this. what I would normally do instead is to look at the posterior samples in a bayesian model to compare each condition and see if they are really different, for $Q, C > N$)

		\item know and question, modal: how can we make argument from model output to differences seen in graph ($Q > C, N > M$)

	\end{itemize}

\paragraph{Discussion and Conclusion.} % (fold)
	\begin{itemize}
		\item Main effect of operator: choice of operator has an effect on projectivity
		\item For \emph{be annoyed}, no significant effect is found for \texttt{question} (vs \texttt{negation}), as would be expected based on Karttunen, but we do find unexpected differences between $N > C, M$ 
		\item for \emph{discover}, we find $Q, C > N$, an effect in the opposite direction from expectation based on Karttunen
		\item \emph{know} shows effects that would be incompatible with characterization as factive or semi-factive: $Q > C, N > M$. if semi-factive, we would expect $N > C, Q$, if factive, no difference would be expected
	\end{itemize}
% paragraph method (end)

\pagebreak
\bibliography{projective-content.bib}

>>>>

%% fonts
\setmainfont[Scale=MatchLowercase,Mapping=tex-text, SmallCapsFeatures={Letters=SmallCaps}]{Times New Roman}
\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Times New Roman}

@@

\vspace{-.4\baselineskip}
	\noindent We present experimental evidence that the projectivity of attitude complements varies across different entailment-cancelling operators, and that this effect of entailment-cancelling operator differs between various attitude verb triggers. We find that verbs show patterns of between-operator projection behavior which group in interesting ways, but 
	do not support \posscite{karttunen_observations_1971} long-standing classification of factive vs. semi-factive verbs, which has been assumed throughout the literature \citep[e.g.][]{djarv_cognitive_2018} + OTHERS.

@@

	\vspace{-.4\baselineskip}
	% Previous work on projection showed that it is not a categorial property of lexical triggers \citep{tonhauser_how_2018}, but a gradient one, affected by various contextual factors \citep{simons_what_2010,de_marneffe_did_2012,beaver_questions_2017,degen_prior_2021}. In light of this, we expect that the hetergeneous entailment-cancelling operators in \ref{ex:family} affect projection differentially.
	%
	\citet{karttunen_observations_1971} suggested that the entailment-cancelling operators in \ref{ex:family} affect projection differentially and proposed a generalization distinguishing \emph{factive} verbs (\emph{regret, forget, resent}) vs. \emph{semi-factive} verbs (\emph{discover, realize, see, find out, notice}): Factives always project, while semi-factives always project across negation, but not always in polar questions or conditional antecedents. Although widely assumed, experimental evaluations of this distinction have been little.

	To provide a systematic way of distinguishing these classes, \cite{djarv_cognitive_2018} suggest that they correspond to emotive and cognitive predicates, respectively. In a study assessing the acceptability of affirmative responses to a target utterance while explicitly denying a potentially projective inference \citep[based on][]{cummins_experimental_2012}, they find that this kind of affirmative cancellation is more readily available for emotive than cognitive predicates, suggesting that the main clause content and projective content are logically more independent from each other for emotive predicates. \cite{smith_relationship_2014}, investigating the effect of operator on the projection of various types of projective content, find an effect on ratings of participant surprisal about the inferences in question, suggesting that inferences triggered by \emph{know} and epithets project more from under negation than conditional antecedents whereas non-restrictive relative clauses show the opposite pattern.

	In our work, we analyze measures from a task designed to measure speaker commitment more directly to address the questions: \textbf{(i)} Is projection affected by differences in entailment-canceling environments? \textbf{(ii)} Do these effects vary for different triggers (and in what way)?

@@

\vspace{-\baselineskip}
\paragraph{Method.} \hspace{-1em}
	The study, originally designed to address a different research question, consisted of 12 experiments, all of which manipulated the factor \texttt{verb} for 20 items corresponding to to the content of the complement clause. Experiments 1--3 used polar question embedding, exps. 4--6 used negation, 7--9 modals, and 10--12 conditionals, making the \texttt{operator} manipulation a between-subjects factor. We analyzed data from 2682 self-identified native English speakers participated online across the 12 experiments (recruited via Prolific and Amazon MTurk). Participants saw items Latin-squared and randomized with six control stimuli.

@@

\vspace{-\baselineskip}
\paragraph{Results and Discussion.} \hspace{-1em}
	Pooling the data across our 12 experiments, we examined the effect on certainty-ratings of \texttt{operator}, and \texttt{verb}, as well as their interaction. Mean certainty-ratings by operator and predicate, and 95\% bootstrapped confidence intervals are shown in  \textbf{Figure \ref{fig:figure1}}.

	% \vspace{-.3\baselineskip}
	% \textbf{version w baseline know/m}
	% \noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{know} and \texttt{modal} as reference levels, and random intercepts for participants and items. We found highly significant main effects of \texttt{operator} (\texttt{conditional}: $+0.142$, \texttt{negation}: $+0.143$, ; \texttt{question}: $+0.225$, where $p < 2e^{-16}$ in all three cases), as well as many interactions of \texttt{operator} and  \texttt{verb} across the board (where $p < 0.001$ in $38$ cases, $p < 0.01$ in one, and $p < 0.05$ in three out of $57$ possible interactions).\\

	\noindent The data was analyzed using a mixed effects linear regression (using \texttt{lme4, lmertest} in \texttt{R}; \citealp{bates_fitting_2015,kuznetsova_lmertest_2016,r_core_team_r_2014}), with \texttt{be\_annoyed} and \texttt{negation} as reference levels, and random intercepts for participants and items.
	% The mean for this baseline (intercept) is $0.867$.
	We found highly significant main effects of \texttt{operator}: For our baseline \texttt{be\_annoyed}, both \texttt{conditional} and \texttt{modal} are clearly less projective than negation, thereby supporting the claim that the embedding context does matter. We also found many interactions of \texttt{operator} and \texttt{verb} across the board, suggesting that the effect of embedding context diffes by verb. Notably, \emph{\lq discover\rq} is more projective in polar questions and conditional antecedents than under negation, patterning opposite to Karttunen's claims about semi-factives. For the emotive predicate \emph{\lq be annoyed\rq} no significant effect is found for \texttt{question} (vs \texttt{negation}), as would be expected based on Karttunen, but we do find unexpected differences between \texttt{negation} $>$ \texttt{conditional, modal}. \emph{know} shows effects that would be incompatible with a characterization as either factive or semi-factive: $\texttt{question} > \texttt{conditional, negation} > \texttt{modal}$. If \emph{\lq know\rq} is a factive predicate, no difference would be expected. If it is semi-factive, we would, again, expect higher projectivity under negation than in questions and conditionals.

	% for \texttt{conditional} ($-0.116$, $p < 1.6e^{-13}$) and \texttt{modal} ($-0.156$, $p < 2e^{-16}$), while the effect of \texttt{question} was only marginally significant ($+0.025$, $p < 0.1$). We also found many interactions of \texttt{operator} and \texttt{verb} across the board (where $p < 0.001$ in $43$ cases, $p < 0.01$ in three, and $p < 0.05$ in one out of $57$ possible interactions).\\

% paragraph results (end)

\vspace{-\baselineskip}
\paragraph{Verb profiles.} % (fold)
	add discussion about verb classes here
% paragraph profiles (end)

\vspace{-\baselineskip}
\paragraph{Conclusion.} % (fold)
	add conclusion here
% paragraph conclusion (end)

\pagebreak
\begin{figure}[t]
		\vspace{-.8\baselineskip}
		\centering
		\includegraphics[width=\textwidth]{graphs/proj-by-both.pdf}\vspace{-1.2\baselineskip}
		\caption{\small Mean certainty ratings by predicate and operator with 95\% bootstrapped confidence intervals (y-axis), by verb (x-axis), and operator (color/grouping, where: \texttt{N} (blue): negation, \texttt{M} (green): modals, \texttt{C} (red): conditional antecedents, \texttt{Q} (purple): polar questions).}
		\label{fig:figure1}
	\end{figure}

\bibliography{projective-content.bib}

### 2022-12-06 materials-lisa/analysis/3_projection-across-operators/data-analysis.Rnw
abstract updates 12/6

<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".6\\textwidth">>=
# packages for data manipulation and plots
require(ggplot2)
require(dplyr)
require(forcats)

@@

pmeans_op %>% mutate(op = fct_reorder(op, Mean,
                                          .fun = mean)) %>%
  ggplot(aes(x = op, y=Mean)) +
  geom_violin(data=data.frame(op=df$op, Mean=df$projective),scale="width",color="gray80") +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue") +
  # labs(title = "Projectivity by operator")+
  theme_bw()

@@

<<summary-verb, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=6>>=

@@

\subsection{Distributions, means, and CIs for projectivity ratings by verb and operator:}
<<summary-combined-1>>=
pmeans = df %>% group_by(verb, op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

ggplot(pmeans, aes(x=verb, y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width = 0.5, color="black") +
  geom_point(size=0.5,color="black") +
  facet_wrap(~op, labeller = labeller(op = c("m" = "Modal", "n" = "Negation", "q" = "Question", "c" = "Conditional"))) +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  ylab("Mean certainty rating") +
  xlab("Predicate")
@

\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
<<summary-combined-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=

pmeans %>% mutate(verb = fct_reorder(verb, Mean,
                                         .fun = mean)) %>%
  mutate(op = fct_reorder(op, Mean, .fun = mean)) %>%
  ggplot(aes(x=verb, y=Mean, group = op, color = op)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(aes(shape = op), size = 3) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
  geom_line() +
  scale_y_continuous(limits = c(0,1)) +
  ylab("") +
  xlab("") +
  guides(shape = "none") +
  scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) +
  labs(title = "", color = "Operator")

>>>>

<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".5\\textwidth">>=
# packages for data manipulation and plots
require(ggplot2)
require(dplyr)
require(forcats)

@@

pmeans_op %>% mutate(op = fct_reorder(op, Mean,
                                          .fun = mean)) %>%
  ggplot(aes(x = op, y=Mean)) +
  geom_violin(data=data.frame(op=df$op, Mean=df$projective),scale="width",color="gray80") +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue")
  # labs(title = "Projectivity by operator")

@@

<<summary-verb, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=4.5>>=

@@

\newpage
\subsection{Distributions, means, and CIs for projectivity ratings by verb, for each operator:}
<<summary-combined-1, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=6>>=
pmeans = df %>% group_by(verb, op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

pmeans %>%
  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective, op=df$op),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width = 0.5, color="black") +
  geom_point(size=0.5,color="black") +
  facet_wrap(~op, labeller = labeller(op = c("m" = "Modal", "n" = "Negation", "q" = "Question", "c" = "Conditional"))) +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  ylab("") +
  xlab("")
@

\newpage
\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
<<summary-combined-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=9, fig.height=4.7>>=

pmeans %>% 
  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean, group = op, color = op)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(aes(shape = op), size = 3) + 
  scale_shape_manual(values = c("C", "M", "N", "Q")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
  geom_line() +
  scale_y_continuous(limits = c(0,1)) +
  ylab("") +
  xlab("") +
  # guides(shape = "none") +
  # scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1))

@
\begin{itemize}
  \item We see interactions between verb and operator

  \item Two verbs show highest projectivity under negation: the anti-veridical \emph{pretend}, and the non-veridical \emph{think}. These are verb with relatively low overall projectivity.

  \item More projective verbs (\emph{\lq announce\rq} and above) have $\texttt{C} > \texttt{M}$

  \item Highly projective verbs (\emph{\lq hear \rq} and above) have  $\texttt{N} > \texttt{M}$

  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
  \begin{itemize}
    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive. The pattern suggested by karrtunen is also not found here (if anything, it might be the other way round)
  \end{itemize}

  \item For more generalizations, let's look at the same information plotted differently: By-operator projectivity for each verb
\end{itemize}

\newpage
\subsection{Projectivity ratings by operator, for each verb (Verb profiles)}
Version 1: With distributions of projectivity ratings as violin-plot
<<verb-profiles-1, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
pmeans %>%
  ggplot(aes(x = fct_reorder(op, Mean), y = Mean)) +
  coord_cartesian(ylim = c(0,1)) +
  facet_wrap(vars(fct_reorder(verb, Mean))) +
  geom_point(size = 1, color = "lightblue") +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective, op=df$op),scale="width",color="gray80") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue") +
  xlab("") + ylab("") +
  labs(title = "Mean projectivity by operator, for each verb")
@
Some more generalizations: We find different \lq profiles\rq\ for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:

\newpage
\noindent Version 2: With lines to show profiles
<<verb-profiles-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
pmeans %>%
  ggplot(aes(x = fct_reorder(op, Mean), y = Mean, group = verb)) +
  coord_cartesian(ylim = c(0,1)) +
  facet_wrap(vars(fct_reorder(verb, Mean))) +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue") +
  xlab("") + ylab("") +
  theme_bw() +
  labs(title = "Mean projectivity by operator, for each verb")
@

\begin{itemize}
  \item \emph{pretend, think}: \textbf{anti-veridical profile}\\
    \texttt{N} $>$ \texttt{M, Q, C}, overall low projectivity

  \item \emph{acknowledge, hear, inform, see, discover, know, be annoyed}: \textbf{\lq factive\rq\ profile}\\
    \texttt{Q} $>$ \texttt{N, C} \textcolor{gray}{$>_?$} \texttt{M}, overall high projectivity (it may be possible to find further subgroups here)

  \item \emph{prove, confirm, establish, demonstrate, (announce), confess, admit, reveal}: \textbf{veridical profile}\\
    \texttt{M, C} $>$ \texttt{Q} $>$ \texttt{N}, overall med-lo to med-hi projectivity

  \item \emph{be right, suggest}: \textbf{reportative profile}\\
    \texttt{M, N, C} $>$ \texttt{Q}

\end{itemize}

Maybe these can have better names, not trying to suggest that verbs can neatly divided in factive v non-factive, but potentially this class / profile is what prompted intuitions in previous literature, and naming in this tradition could make sense, but can be changed depending on our rhetoric, of course.

\newpage

<<verb-profiles-3, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
vgroups <- data.frame(verb = levels(pmeans$verb), profile = 
                       c("0+0",  # acknowledge
                         "00+",   # admit 
                         "00+",  # announce
                         "-+-",  # be annoyed
                         "+-+",    # be right
                         "00+",   # confess
                         "+-+",   # confirm
                         "+0+",   # demonstrate
                         "0+0",  # discover
                         "+-+",   # establish
                         "-+0",  # hear
                         "-+-",  # inform
                         "-+-",  # know
                         "-00",    # pretend
                         "+-+",   # prove
                         "0++",   # reveal
                         "0-+",    # say
                         "-+-",  # see
                         "0-+",   # suggest
                         "-00"     # think
                         ))

lookup_group <- function(verb_name){
  return(as.character(filter(vgroups, verb == verb_name)[2]))
}

pmeans = pmeans %>% rowwise() %>% mutate(groups = lookup_group(verb))

pmeans %>%
  ggplot(aes(x = fct_reorder(op, Mean), y = Mean, group = verb, color = verb)) +
  coord_cartesian(ylim = c(0,1)) +
  facet_wrap(~groups) +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1) +
  geom_line() +
  xlab("") + ylab("") +
  theme_bw() +
  labs(title = "Mean projectivity by operator, for each verb")
@

### 2022-12-06 materials-lisa/analysis/3_projection-across-operators/data-analysis.tex
abstract updates 12/6

{\centering \includegraphics[width=.6\textwidth]{figures/summary-op-1}

@@

\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\linewidth]{figures/summary-combined-1}

>>>>

{\centering \includegraphics[width=.5\textwidth]{figures/summary-op-1}

@@

\newpage
\subsection{Distributions, means, and CIs for projectivity ratings by verb, for each operator:}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\linewidth]{figures/summary-combined-1-1}

\newpage
\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\linewidth]{figures/summary-combined-2-1}

\end{knitrout}
\begin{itemize}
  \item We see interactions between verb and operator

  \item Two verbs show highest projectivity under negation: the anti-veridical \emph{pretend}, and the non-veridical \emph{think}. These are verb with relatively low overall projectivity.

  \item More projective verbs (\emph{\lq announce\rq} and above) have $\texttt{C} > \texttt{M}$

  \item Highly projective verbs (\emph{\lq hear \rq} and above) have  $\texttt{N} > \texttt{M}$

  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
  \begin{itemize}
    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive. The pattern suggested by karrtunen is also not found here (if anything, it might be the other way round)
  \end{itemize}

  \item For more generalizations, let's look at the same information plotted differently: By-operator projectivity for each verb
\end{itemize}

\newpage
\subsection{Projectivity ratings by operator, for each verb (Verb profiles)}
Version 1: With distributions of projectivity ratings as violin-plot
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\linewidth]{figures/verb-profiles-1-1}

\end{knitrout}
Some more generalizations: We find different \lq profiles\rq\ for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:

\newpage
\noindent Version 2: With lines to show profiles
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}

{\centering \includegraphics[width=\linewidth]{figures/verb-profiles-2-1}

\begin{itemize}
  \item \emph{pretend, think}: \textbf{anti-veridical profile}\\
    \texttt{N} $>$ \texttt{M, Q, C}, overall low projectivity

  \item \emph{acknowledge, hear, inform, see, discover, know, be annoyed}: \textbf{\lq factive\rq\ profile}\\
    \texttt{Q} $>$ \texttt{N, C} \textcolor{gray}{$>_?$} \texttt{M}, overall high projectivity (it may be possible to find further subgroups here)

  \item \emph{prove, confirm, establish, demonstrate, (announce), confess, admit, reveal}: \textbf{veridical profile}\\
    \texttt{M, C} $>$ \texttt{Q} $>$ \texttt{N}, overall med-lo to med-hi projectivity

  \item \emph{be right, suggest}: \textbf{reportative profile}\\
    \texttt{M, N, C} $>$ \texttt{Q}

\end{itemize}

Maybe these can have better names, not trying to suggest that verbs can neatly divided in factive v non-factive, but potentially this class / profile is what prompted intuitions in previous literature, and naming in this tradition could make sense, but can be changed depending on our rhetoric, of course.

\newpage

{\centering \includegraphics[width=\linewidth]{figures/verb-profiles-3-1}
